---
title: "README"
author: "Ellen Esch"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  github_document:
    toc: yes
always_allow_html: yes
urlcolor: blue
css: "./inst/app/www/stye.css"
---

```{r setup, include=FALSE, eval = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = F, warning = F)

```

## Overview

Growing Shade is a prioritization tool for tree canopy enhancement and preservation. Growing Shade was conceptualized by the Metropolitan Council, Tree Trust, and The Nature Conservancy with a focus on the Twin Cities (Minnesota) area. All analyses and code were developed by Ellen Esch. The Twin Cities version of Growing Shade can  be accessed at: www.growingshade.com or https://metrotransitmn.shinyapps.io/growing-shade/. 

The components of Growing Shade relevant to population demographics (i.e., determining environmental justice priority areas) can be scaled to any area within the United States. The components relevant to the tree canopy have been parameterized to the Twin Cities region, but should work pretty well in other temperate areas with deciduous and coniferous trees.

**This repository walks through the creation and aggregation of data used in interactive mapping tool.**

## Getting started

The `Tutorial.Rmd` script walks through the steps to get a working versions for different areas. 

Most packages are available on CRAN, so can be downloaded through standard procedures. There are two specific packages which should be downloaded from GitHub versions.

```{r}
devtools::install_github("eclarke/ggbeeswarm")
devtools::install_github("Metropolitan-Council/councilR")
```
