---
title: "Growing shade report"
date: "`r format(Sys.time(), '%d %B %Y')`"

params:
  param_area: NA
  param_min: NA
  param_max: NA
  param_ntracts: NA
urlcolor: blue
---

<style type="text/css">
.caption {
    font-size: .9em;
    font-color: #A9A9A9
}
</style>


```{r assignments, echo=F, message=F, warning=F, results='hide'}
library(dplyr)
library(stringr)
library(ggplot2)
library(tidyr)

knitr::opts_chunk$set(echo = TRUE, fig.align="center")

```

# Custom report for `r params$param_area`

## Tree canopy summary

 ```r params$param_area``` has an existing tree canopy which ranges from ```r params$param_min```% to ```r params$param_max```% across ```r params$param_ntracts``` Census tracts. The distribution of tree canopy across the region is shown below; tracts in ```r params$param_area``` are highlighted in green.

In most areas in our region, a tree canopy coverage of 40% (as detected by our methods) leads to the greatest benefits. Note that native tallgrass prairie occurs throughout our region - lower tree coverage in areas dominated by tallgrass prairie should not be penalized.

```{r}


```

**This is a work in progress. When the in-application report is finalized, it will also be finalized here. Thanks for your patience!**
