---
title: "Growing shade report"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 3
params:
  selected_tract: NA
  selected_geo: NA
  selected_city: NA
urlcolor: blue
---

```{r echo=F, message=F, warning=F, results='hide'}
#This is the selected tract.
params$selected_tract
params$selected_geo
params$selected_city

reportfor <- if (params$selected_geo == "City (MetCouncil region only)") {
  params$selected_city
} else if (params$selected_geo == "Selected tract") {
  paste0("tract ", params$selected_tract)
} else if (params$selected_geo == "County (within MN or WI)") {
  'PRANK! nothing to see here yet'
}
```


# Report for ```r reportfor```.

This report might be super cool eventually!

## Status of the tree canopy

Understand how the tree canopy of the selected region compares to other geographies across the region.

```{r status, echo=F, message=F, warning=F}
library(tidyverse)
ggplot(mtcars, aes(x = mpg, y = cyl)) +
  geom_point()
```

## Equity analysis

Understand how people and trees interact, and where the selected geography compares.

## Trees past and present (with lessons for the future)

Something about tree community composition. 
