---
title: "Growing Shade Use Report"
# author: "Ellen"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  bookdown::word_document2:
    reference_docx: counciltemplate.docx
    toc: true
    plots:
      caption:
        tnd: 1
    tables:
      caption:
        tnd: 1
      conditional:
        first_row: true
        first_column: true
        last_row: false
        last_column: false
        no_hband: false
        no_vband: true
urlcolor: blue
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = F, message = F, warning = F, results = "hide"#,
  # fig.pos = "H"
  # ,fig.height = 4
  , dpi = 300
)


library(tidyverse)
library(flextable)
load("../data/ctu_list.rda")
load("../data/metc_region.rda")

df <- read_csv("./analytics-export.csv", skip = 312) %>%
  filter(!str_detect(City, "Cheyenne|not set"),
         Region == "Minnesota") %>%
    mutate(City = case_when(City == "Saint Paul" | City =="South Saint Paul" ~ "St. Paul",
                            City == "West Saint Paul" ~ "West St. Paul",
                            TRUE ~ City)) %>%
  group_by(City) %>%
  summarise(Users = sum(Users),
            `Event count` = sum(`Event count`),
            `Engagement time` = sum(`Average engagement time` * Users))  %>%
  mutate(`Average time engaged (min)` = `Engagement time` / Users / 60) %>%
  left_join(ctu_list, by = c("City" = "GEO_NAME")) %>%
  sf::st_as_sf()
```

# Introduction 

**Growing Shade** is a tree canopy enhancement and preservation tool which combines *local stories* and an *interactive mapping tool*.

The groundwork for Growing Shade was laid with Thrive MSP 2040 which included outcomes of *sustainability* and *equity.* A partnership with The Nature Conservancy and Tree Trust was formed and a tree canopy tool was conceptualized in 2020(?). An advisory group was formed and a stakeholder survey was conducted in early 2021 to inform the project. Growing Shade was developed over the summer and fall of 2021. The project had a soft launch in November 2021 for the purposes of soliciting user feedback and beta testing the project. Growing Shade was refined in response to the feedback, and official launched late January 2022. 

This summary is intended to inform how successful (or not) outreach efforts have been thus far, and may be useful in refining future efforts and continuing to promote usage of the application to inform a more sustainable and equitable future for the Twin Cities region. 

## Presentations conducted

Starting in January 2022, Growing Shade has been presented to practitioners with a focus on foresters and planners. 

- Committee of the Whole (Feb 2, 2022)
- MN Shade Tree Advisory Committee (Feb 17, 2022; ~50 attendees)
- PlanIt Implementation Webinar (Feb 24, 2022; 70 registrants)
- MN Shade Tree Short Course (March 15, 2022; 106 participants attended the session)
- Washington County Water Consortium (May 4, 2022)

## News coverage

Starting in April 2022, the focus of outreach efforts switched to general audiences.

- [Shady practices: New mapping tool shows inequity of tree coverage in the metro](https://www.minnpost.com/environment/2022/04/shady-practices-new-mapping-tool-shows-inequity-of-tree-coverage-in-the-metro/). MinnPost. April 29, 2022.
- [A new tool lets metro residents track region's tree canopy](https://www.mprnews.org/episode/2022/04/29/a-new-tool-lets-metro-residents-track-regions-tree-canopy). Minnesota Public Radio. April 29, 2022. 
- [Met Council, cities team up to giveaway trees to plant in your yard](https://www.kare11.com/article/news/local/met-council-cities-team-up-to-giveaway-trees-to-plant-in-your-yard/89-124fa8ed-7075-41d0-b3b7-47141f769b0c). Kare 11. April 29, 2022.

\newpage

## Application usage

Since January 2022, Growing Shade has been accessed by 581 users (929 users since November 2021). This summary only looks at user data since January 2022. We don't know exactly how accurate internet provider locations are (and some people use vpns so we can't access the information), but this information may still be valuable! Note, all development traffic is run locally, so should not be contaminating these statistics. 

### Who is using the app?

Within the 7 county region, St. Paul and Minneapolis have the most users (Figure \@ref(fig:user-fig)). 

```{r user-fig, fig.cap='Users accessing the Growing Shade application by city.', out.width='90%', fig.asp=.5, fig.align='center'}
df %>% 
  # arrange(-Users) %>%
  # select(City, Users) %>%
  ggplot() +
  geom_sf(data = metc_region) +
  geom_sf(aes(fill = Users), lwd = .5) +
  theme_void()

```

\newpage

But the east metro (esp. Maplewood) has the highest engagement where people have spent the most time, on average, within the application (Figure \@ref(fig:time-fig)).

```{r time-fig, fig.cap='Average time that each user has spent engaged on the Growing Shade application.', out.width='90%', fig.asp=.5, fig.align='center'}

df %>% 
  # mutate(EpU = `Event count` / Users) %>%
  ggplot() +
  geom_sf(data = metc_region) +
  geom_sf(aes(fill = `Average time engaged (min)`), lwd = .5) +
  theme_void()+
  labs(fill = "Average\ntime\nengaged\n(min)")
```

While Growing Shade has a scope limited to the 7 county region, we do see engagement across the state (and also in neighboring states). Table \@ref(tab:other-city-table) lists the other cities in Minnesota where more than one person has accessed Growing Shade. 

```{r, other-city-table, results='markup', tidy=FALSE}
regulartable(
  df %>%
  filter(is.na(ntracts),
         Users > 1) %>%
  arrange(-Users) %>%
  sf::st_drop_geometry() %>%
  select(City, Users)
) %>%
  theme_zebra() %>%
  # theme_booktabs() %>%
  autofit(add_w = 0, add_h = 0) %>%
  set_table_properties(layout = "autofit") %>%
  set_caption("Other Minnesota cities where more than one person has accessed Growing Shade since January 2022.")  %>%
    fontsize(size = 10, part = "all")

```

### How are people using the app?

Most people are scrolling and clicking around the application. Since January 2022, there have been 30 file downloads. 

### How are pepole finding the app?

Since January 2022, new users have mostly accessed the application via the direct link (435 users; https://metrotransitmn.shinyapps.io/growing-shade/). A growing number of users (160) are accessing the application via a referral from metrocouncil.org. Other ways of accessing the application (socials like linkedin, facebook, youtube, teams are lower - it's unclear if Twitter gets pulled out separately or if those users are part of the "direct link" users).

Just under 6% of users are accessing the application via a mobile platform. 
